Self-tunable approximated explicit MPC: Heat exchanger implementation and analysys

### 2023-02-03 paper/elsarticle/main.tex
Review: Conclusions

	This paper deals with the experimental implementation of the self-tunable approximated explicit model predictive control and provides a strategy for an effective self-tuning controller design. Based on the current reference value, the tuning parameter is scaled using linear interpolation. The previously published work related to the self-tunable explicit MPC suggested tuning based on the distance of the reference value from the system steady-state value. This paper presents a novel idea of tuning based on the size of reference step change. The self-tuning algorithm aims to compensate for the nonlinear behavior of the controlled system. The self-tuning parameter is updated when the reference changes. It is calculated as the ratio between the size of the reference change and the maximal possible size of the reference change, which is specified before operation. 
	Another novel contribution addresses the challenging control problem of asymmetric system behavior by splitting the interval of tuning parameter into two ranges, while both intervals are assigned to different operating conditions. The proposed method is implemented on a laboratory heat exchanger with nonlinear and asymmetric behavior. The asymmetry makes the plant a suitable candidate for splitting the interval of the tuning parameter. The decision criterium is negativity or positivity of reference change. When the reference changed upwards, the control input was tuned in the first part of the interval and approached the boundary controller associated with the lower bound on the penalty matrix. On the contrary, when the reference changed downwards, the control input was tuned to approach the control input from the boundary controller with the upper bound on the penalty matrix. To investigate the control results properly, the control performance was evaluated also quantitatively. Compared to the conventional control strategy handling just a single controller, the self-tunable control approach decreased the integral squared control error, maximal overshoots/undershoots, and settling time. In addition, the energy consumption corresponding to every controller was evaluated as well. The value of consumed energy was between the two boundary controllers, i.e., the energy consumption did not deteriorate in general. With saving energy compared to one of the boundary controllers, the control performance was significantly improved.

>>>>

	This paper deals with the experimental implementation and analysis of the novel self-tunable approximated explicit model predictive control method and provides a strategy for an effective self-tuning controller design. Based on the current value of the piece-wise constant reference, the tuning parameter is scaled using linear interpolation. 
	In contrast to the previously published work related to the self-tunable explicit MPC suggested tuning based on the distance of the reference value from the system steady-state value. This paper presents a novel perspective idea of the self-tuning based on the size of reference step change. The self-tuning algorithm aims to compensate for the nonlinear behavior of the controlled system. The self-tuning parameter is updated whenever the reference changes. the tuning value is calculated as the ratio between the size of the reference change and the maximal admissible size of the reference change, which is specified before operation. 
	Another novel contribution addresses the challenging control problem of asymmetric system behavior by splitting the interval of the self-tuning parameter into two ranges, while both intervals are assigned to different operating conditions. The proposed method is implemented on a laboratory-scaled heat exchanger with nonlinear and asymmetric behavior. The asymmetry makes the plant a suitable candidate to analyze the benefits of splitting the interval of the tuning parameter. The decision criterium is negativity or positivity of reference change. When the reference changed upwards, the control input was tuned in the first part of the interval and approached the boundary controller associated with the lower bound on the selected penalty matrix. On the contrary, when the reference changed downwards, the control input was tuned to approach the control input from the boundary controller with the upper bound on the penalty matrix. To investigate the control results properly, the control performance was judged also quantitatively using a set of quality criteria. Compared to the self-tunable control approach outperformed the conventional control strategy handling just a single controller, i.e., non-tunable controller, by minimizing the integral squared control error up to 32\,\%, reducing maximal overshoots/undershoots up to 18\,\%, and decreasing settling time up to 24\,\%. In addition, the energy consumption corresponding to each of the constructed controllers was also analyzed. The energy consumption of the self-tunable explicit MPC controller was in between the values of the two boundary non-tunable explicit MPC controllers, i.e., the total energy consumption trades off the performance of the boundary controllers. % did not deteriorate in general. 
	With saving the energy compared to one of the boundary controllers, the control performance was significantly improved.
